---
output: 
  github_document:
    pandoc_args: --webtex
bibliography: doc/used_cars.bib
---

# Used Cars Price Prediction
DSCI-522 - Group 308
Authors: Andrés Pitta, Braden Tam, Serhiy Pokrovskyy

## About

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(tidyverse)
library(caret)
```

```{r load data, echo=FALSE, message=FALSE, warning=FALSE}
train_scores_text <- read_csv("results/train_metrics.csv") %>%
  dplyr::rename(train = 'Train Score',
                valid = 'Validation Score')
test_scores_text <- read_csv("results/test_metrics.csv") %>%
  dplyr::rename(value = 'Metric Value')
```

In this project we attempt to build a regression model which can predict the price of used cars based on numerous features of the car. We tested the following models: support vector regression, stochastic gradient descent regression, linear regression, K-nearest neighbour regression, and random forest regression. data set is not linearly separable, more clustered. We found that support vector regression had the best results, having an $R^2$ score of `r round(train_scores_text$train[1],3)` on the training set, $R^2$ score of `r round(train_scores_text$valid[1],3)` on the validation set and $R^2$ score of `r round(test_scores_text$value[1],3)` on the test set. Given that the dataset was imbalanced, this led to poor prediction of the classes that were quite sparse.

The data set used in this project is Used Cars Dataset created by Austin Reese. It was collected from Kaggle.com [@reese_2020] and can be found [here](https://www.kaggle.com/austinreese/craigslist-carstrucks-data). This data consists of used car listings in the US scraped from Craigslist that contains information such as listed price, manufacturer, model, listed condition, fuel type, odometer, type of car, and which state it's being sold in.

## Report
The final report can be found [here](https://github.com/UBC-MDS/DSCI_522_Group-308_Used-Cars/blob/master/doc/used_cars_report.md).

## Usage
To replicate the analysis, clone this GitHub repository, install the dependencies listed below and follow these instructions:

Before starting, if you want to reset all the outputs, run the following command at the command line/terminal:

    make clean

**Warning** The dataset's size is 1.35GB. In order to replicate a quick version of the model with 1% of the data (see makefile documentation to change train size), run the following command at the command line/terminal:
  
    make quick
  
If you want to run the model with all the data run:
  
    make all
    
## Dependencies
- Python 3.7.3 and Python packages:
    - altair==3.2.0
    - selenium==3.141.0
    - docopt==0.6.2
    - pandas==0.24.2
    - numpy==1.16.4
    - statsmodel==0.10.0
    - plotly==4.3.0
    - scikit-learn==0.20.4
- R version 3.6.1 and R packages:
    - knitr==1.24
    - docopt==0.6.1
    - tidyverse==1.3.0
    - readr==1.3.1

## Notes

Please be advised that in order to reproduce the analysis, the `scripts/download.py` script will download approximately 1.4GB original data file. Also, training the model with `scripts/train_model.py` may take several hours. You may consider using command line arguments to train on a configurable subset of training data (which may affect the trained model) Please see consult usage data on each script for further details and options.
  
## License
The Used Cars Prediction materials here are licensed under the Creative Commons Attribution 2.5 Canada License (CC BY 2.5 CA). If re-using/re-mixing please provide attribution and link to this webpage. 

# References